On The Impact of Atoms of Confusion in JavaScript Code

### 2022-07-25 3_Method.tex
Finished reviewing the methodology section. Next is the results section.

To complement the two experiments, we perform semi-structured interviews with professional JavaScript developers, aiming to identify their perceptions regarding programs containing atoms of confusion (research question RQ2). We also ask each participant if they know of any other JavaScript-specific construct or idiom that they regard as being likely to make the code hard to understand. % (research question RQ3). 
 In this section we detail the protocol we followed to conduct the interviews and to analyze the results.

@@

In the third part of the interview, participants are shown pairs of programs that were used in the first study, where each pair contains a clean and an obfuscated version of the same atom candidate. The participants 
%were told both sides of each pair had the same output, and 
are asked to evaluate which version of the code is easier to understand. To avoid introducing bias in the answers, the interviewers do not explain that one of the versions in each pair contains the atom under investigation. Subjects are just presented the pairs of programs and allowed to take the necessary time to decide on the most readable one.

@@

To understand how often the analyzed atom candidates appear in open source projects, and thus answer our third research question (\emph{What is the frequency of occurrence of atoms of confusion in practice?}), we mined a set of GitHub open source repositories. To this end, we first collected the most popular GitHub repositories that are primarily written in JavaScript. We measured popularity using the project's stargazers. This metric, available through the GitHub API, represents the number of stars a project received from users of the platform. The same metric has been used in a number of previous studies as a proxy to estimate a project's popularity~\cite{gyimesi2019bugsjs,canedo:esem2020}. We then selected the top 100 most popular repositories and removed projects that did not reach the first quartile of the distribution of lines of code.

After filtering out JavaScript project candidates, in the second step we built a curated dataset comprising the top \minedprojects repositories. Examples of projects in this dataset include \textsc{React}, \textsc{Node JS}, and \textsc{AngularJS}. Table~\ref{tab:projects-statistics} presents some statistics about the projects we consider in our research. The size of the projects range from small ones (5543 lines of code) to complex systems with more than 1 MLOC. All projects in our dataset have at least \num{1244} forks and at least \num{23672} stars. We automated all the steps to filter, clone, and collect the statistics from the repositories using Python scripts.

We mined the occurrence of atom candidates from the repositories in our curated dataset using source code queries that we wrote using the CodeQL language~\cite{moor:gttse2007}. CodeQL is an object-oriented variant of the Datalog language~\cite{rodriguez2020efficient}, and currently supports researchers and practitioners to query the source code of systems written in different languages (such as \cpplang, Java, and JavaScript). We also automate the process of running the queries and exporting the results to a format that simplifies our analysis (and also the reproduction of this study). Finally, we computed some descriptive statistics to measure the prevalence of atoms of confusion in practice.

>>>>

To complement the two experiments, we perform semi-structured interviews with professional JavaScript developers, aiming to identify their perceptions regarding programs containing atoms of confusion (research question RQ2). We also ask each participant if they know of any other JavaScript-specific construct or idiom they think is likely to make the code hard to understand. % (research question RQ3). 
In this section we detail the protocol we followed to conduct the interviews and to analyze the results.

@@

In the third part of the interview, participants are shown pairs of programs that were used in the experiments, where each pair contains a clean and an obfuscated version of the same atom candidate. For this part we only use atoms that appear in both experiments. The participants 
%were told both sides of each pair had the same output, and 
are asked to evaluate which version of the code is easier to understand. To avoid introducing bias in the answers, the interviewers do not explain that one of the versions in each pair contains the atom under investigation. Subjects are just presented the pairs of programs and allowed to take the necessary time to decide on the most readable one.

@@

To understand how often the analyzed atom candidates appear in open source projects, and thus answer our third research question (\emph{What is the frequency of occurrence of atoms of confusion in practice?}), we mine a set of GitHub open source repositories. To this end, we first collect the most popular GitHub repositories that are primarily written in JavaScript. We measure popularity using the project's stargazers. This metric, available through the GitHub API, represents the number of stars a project received from users of the platform. The same metric has been used in a number of previous studies as a proxy to estimate a project's popularity~\cite{gyimesi2019bugsjs,canedo:esem2020}. We then select the top 100 most popular repositories and remove projects that do not reach the first quartile of the distribution of lines of code.

After filtering out JavaScript project candidates, in the second step we build a curated dataset comprising the top \minedprojects repositories. Examples of projects in this dataset include \textsc{React}, \textsc{Node JS}, and \textsc{AngularJS}. Table~\ref{tab:projects-statistics} presents some statistics about the projects we consider in our research. The size of the projects range from small ones (5543 lines of code) to complex systems with more than 1 MLOC. All projects in our dataset have at least \num{1244} forks and at least \num{23672} stars. We automate all the steps to filter, clone, and collect the statistics from the repositories using Python scripts.

We mine the occurrence of atom candidates from the repositories in our curated dataset using source code queries that we wrote using the CodeQL language~\cite{moor:gttse2007}. CodeQL is an object-oriented variant of the Datalog language~\cite{rodriguez2020efficient}, and currently supports researchers and practitioners to query the source code of systems written in different languages (such as \cpplang, Java, and JavaScript). We also automate the process of running the queries and exporting the results to a format that simplifies our analysis (and also the reproduction of this study). Finally, we compute some descriptive statistics to measure the prevalence of atoms of confusion in practice.
